---
title: "Lauren Lifke"
output: html_document
date: "2025-04-22"
editor_options: 
  markdown: 
    wrap: 72
---

**#rsw comment: Lauren - Great to see you hand in a rendered html
document! Progress. I went to your repo to edit the source document and
put comments in the text throughout. You have come a long way in
focusing this topic, so great job. The next steps will involve adding
some data visualizations and some tables to your text and integrating
the data into your narrative. We will work that today so you are in good
shape.**

Also note some of the formatting changes I made, such as numbered lists
and bullets.

**Rob**

Lifke story pitch: \# Lauren Lifke \#
[lifke.laur\@gmail.com](mailto:lifke.laur@gmail.com){.email} \# June 2

#rsw comment - strong introduction and focus.

During the 2024 election, David D. Smith, CEO of Sinclair Broadcast
Group and owner of the Baltimore Sun, donated the sixth-highest dollar
amount out of all donors to receiving committees, according to
Maryland's 2024 campaign finance data. The donations, totaling \$85,000
and made through two transactions in October 2024, were made out to
People for Elected Accountability and Civic Engagement -- a ballot issue
committee that aimed to shrink the size of the Baltimore City Council.
Though that ballot measure didn't pass in November, Smith's influence on
the Sun's coverage remains a hot topic more than a year after its
purchase, especially as the topics of press freedom and political
influence on media have grown nationally since President Donald Trump's
inauguration.

Baltimore citizens are the target audience, as they voted against the
ballot measure 62-38%, **rsw: 62%-38%** in part with the help of Mayor
Brandon Scott and the Stop Sinclair Committee.

Interview subjects:

1.  A spokesperson for PEACE who can comment on the donations

2.   A spokesperson for Stop Sinclair, the committee that succeded in
    its goal of not allowing the ballot measure to pass

3.  Current or former reporters for the Baltimore Sun who can speak on
    the shift in culture after Smith's purchase

4.  The data to compare the size of Smith's donations with others

    **Potential impact:** By digging deeper into the specifics of the
    campaign donations, which have already been reported on, the article
    can keep the conversation and momentum going regarding Smith's
    influence on local politics and the media in Baltimore. Discussion
    of this topic has already helped sway voters **\*\*not sure this is
    correct - check! \*\*** [to vote against a ballot measure for the
    first time in 20 years,]{.underline} and keeping the conversation
    moving might allow for change in future elections, too.

**Three surprising or interesting things:** It was surprising to see
that an individual, as opposed to an organization, was in the top 10
contributors. I also learned more about the local Maryland politics just
by discussing the story with classmates and mentors, so it was
interesting to grow more invested in the community I'll be reporting in.
Aside from that, I also learned that, although the contribution amount
caps at an allotted \$6,000 per person, there is an exception for ballot
issue committees. **\*\*rsw comment: Good catch! \*\***

<<<<<<< HEAD
=======
Previous coverage: The Baltimore Banner is a newer newsroom that has
been following Smith's purchase of the Baltimore Sun and his involvement
in the ballot measure, and its coverage, linked below, highlights the
impact that shrinking a city council would have had on smaller
communities. It has also followed the Stop Sinclair committee, showing
that the fight against Smith has grown expensive.

I can make this story visual by comparing the size of Smith's donations
with other individuals and groups. The top 20 donors exceeded other
donations by a large margin, so I can show how Smith compares to them
both in number and by issue using pie charts or bar graphs. I can also
use older data to show how Smith's contributions have changed over time.

**\*\*rsw comment: good to identify data gaps** \*\*I need to gather
newer data -- according to Maryland State Board of Elections,
disclosures of contributions are due biannually, with the most recent
being May 31. These contributions aren't available publicly online yet
from what I can tell, but once it's up I can gather that information and
include it to add to the timeliness aspect of the story.
<https://elections.maryland.gov/campaign_finance/disclosure_of_contributions.html>

I can deliver the aspects of the story that are in my control by
\*\***rsw comment: probably not! these stories take a lot of time**\*\*
[Friday]{.underline}, and hopefully the newer data will be available
before then so that I can I can include that.

**\*\*rsw comment: don't put your notes on a reporting memo. it should
be clean and complete** \*\*

~~##Notes Your Name, Contact Information, Date In two sentences, what is
this story about? In one sentence, why tell this story NOW? Who is your
target audience? Three people you can interview for this story Potential
impact What are three things that surprised or interested you about this
story? Context: Summarize any previous coverage. Photos, video,
graphics - How will you make this story visual? What other information
do you need to gather? Estimated delivery~~

>>>>>>> 5aeece36d897bd6159c145766988cc7431298d32
Related coverage:

<<<<<<< HEAD
### Code
=======
-   <https://www.rollingstone.com/politics/politics-features/baltimore-sun-right-wing-takeover-david-smith-1235268329/>

-   <https://www.thebaltimorebanner.com/politics-power/local-government/baltimore-council-ballot-size-E6DOSA5QD5E3JIRGQHUV53WXUA/>

-   <https://www.wypr.org/wypr-news/2024-11-05/baltimore-city-voters-reject-ballot-question-h-and-keep-city-council-current-size>/

-   <https://www.thebaltimorebanner.com/economy/sun-owner-david-smith-fox-VQKQHEG4EJAIVA6TRKXMKR5EUM/>

-   
    <https://www.thebaltimorebanner.com/politics-power/local-government/baltimore-council-reduction-david-smith-BQDKG67DN5BJ5IOUPIXB4IO3AU/>

**\*\*rsw comment: cut out this stuff. focus on your project**\*\*

~~A demo notebook for Dow Jones News Fund data interns~~

~~During the week’s training, you will work on a project called
Investigate Your Community to produce detailed story pitches relevant to
their summer internship positions. The goal is to have the students
research a topic relevant to your newsroom so they can get to work
immediately at the internship.~~

~~Here are some examples from last year’s DJNF data journalism
training:~~

~~Hannah Rosenberger, at ASI Media/Counselor magazine, produced this
analysis on name, image and likeness trends:
<https://hrosenberger.github.io/ASI-NIL-Project/>~~

~~Shifra Dayak, an intern at the Connecticut Mirror, produced this
analysis of environment spill data:
<https://shifradayak.github.io/DJNF_Merrill/ct_env_spills.html>.~~

~~Michael Zhang, an intern at the Houston Landing, explored local
election data:
<https://michaelzhangreporter.github.io/Houston_Campaign_Finance/>~~

~~Here's what your newsroom suggested for a project Campaign finance
data~~

~~This notebook helps you get started~~

>>>>>>> 5aeece36d897bd6159c145766988cc7431298d32
# Load software libraries

```{r}
#install.packages("tidyverse")
#install.packages("janitor")
#install.packages("readxl")
#install.packages("rvest")
library(tidyverse)
library(janitor)
library(readxl)
library(rvest)

```

# Load your data

```{r}

# 
# YOUR_DATA <- read.csv("../data/XXXXX.csv") |> 
#     clean_names()

data <- read.csv("./data/2024 Presidential Post General Md Financial Info 2024 Presidential Post General .csv") |> 
    clean_names()


```

Data details:

# Explore the data types

```{r}

#glimpse(YOUR_DATA)

glimpse(data)

```

# Fix dates

```{r}
data <- data |> 
  mutate(date = lubridate::mdy(contribution_date))
```

# Data exploration: dates

```{r}
data |> 
   group_by(date, contribution_type) |>
  summarize(count = n(), total =sum(contribution_amount)) |>
  arrange(desc(count))
```

<<<<<<< HEAD
=======
```{r}
MSEA <- data |>
  group_by(receiving_committee) |>
  summarise(num = n(),
          receiving_total = sum(contribution_amount))
```

# Questions

Write down questions you want answered:

-   Has anybody exceeded the \$6,000 limit?

-   Has anybody exceeded the total \$24,000 limit?

-   What day saw the most political contributions?

-   What month saw the most political contributions?

-   What month saw the fewest political contributions?

-   Why was Nov. 8, 2024 one of the dates with the largest number of
    contributions, three days after the election?

-   Which committee received the highest number of contributions?

-   Which committee received the highest total dollar amount in
    contributions?

>>>>>>> 5aeece36d897bd6159c145766988cc7431298d32
# Contributors

```{r}
contributor_total <- data |>
group_by(contributor_name) |>
  summarise(num = n(),
          total = sum(contribution_amount))
```

```{r}
receivingcommittee_total <- data |>
  group_by(receiving_committee) |>
  summarise(num = n(),
          receiving_total = sum(contribution_amount))
```

```{r}
data |> 
   group_by(receiving_committee) |>
  summarize(count = n(), total =sum(contribution_amount)) |>
  arrange(desc(count))
```

<<<<<<< HEAD


=======
# Data Details

Maryland Campaign Reporting Information System\
<https://campaignfinance.maryland.gov/Public/ViewReceiptsMain>.
Financial Information\
2024\
11/19/2024 Post General Presidential period\
Electoral\
1/1/2-24-12/31/2024

Some General Information regarding contributions, transfers and loans.

A contribution is money or anything of value given to a political
committee to promote or assist in promoting the success or defeat of a
candidate, political party, or question. A contribution can be made by
any individual or by an entity, including a corporation or other
business entity, a political club, a federal committee, or a labor
union.

A transfer is a monetary contribution made by one political
committee  to another. For example: a PAC may transfer money to a
candidate committee. In addition, the law allows an out-of-state
political committee to transfer money to a Maryland political committee
if the out-of-state committee is properly registered with its own
state’s election authority. Federal committees are not considered
out-of-state political committees and therefore cannot make transfers
but can make contributions. 

A loan is a permissible way for a campaign to receive funds. There is no
limit on the amount of money that can be loaned to the campaign.
However, unless the loan is executed properly and paid off in a timely
manner, the loan will be converted into a contribution.

A person may contribute directly or indirectly no more than \$6,000 to
one political committee, and a total of \$24,000 to all political
committees, during the four-year cycle.

The \$6,000 limit on contributions to the political committee of a
candidate applies regardless of the number of offices sought by the
candidate in separate elections during the 4 year election cycle or the
number of authorized candidate committees formed to support the
candidate, including slates.

A  political committee may transfer no more than \$6,000 to another
political committee during the four (4) year cycle.  There is no
aggregate limit.

FAQs #VALUE! #VALUE!\
Categories:\
ReConnecting Your Committee\
Committee Registration\
Bank Accounts\
Using Campaign Finance Funds\
Campaign Finance\
General Questions\
Closing Your Committee\
Independent Expenditures and Participating Organizations\
Making Expenditures\
Slates

Question and Answers:\
ReConnecting Your Committee\
Que. Why does the verification show that I am not a registered voter
when I am?\
Ans. That is because the piece of information provided, either your last
4 digits of your Social Security number or Driver's license, could not
be verified against the voter registration records. The best form of ID
is your voter id number. If you don't know that number, please contact
the State Board of Elections at 410-269-2880.

Que. After one officer has activated the committee with MD CRIS, does
the other officer need to as well?\
Ans. Yes. The other officer must go through the same initial process via
ReConnect to Your Committee in order to receive a temporary password.
The email account used during the process will be the username. Once the
other officer has received the email, please log in and change the
password. The other officer will be able verify the filed report under
My verifications.

Committee Registration\
Que. If I send in a new appointment for a Treasurer and/or Chairman, do
I immediately replace the previous officer?\
Ans. No, the previous officer must resign in writing prior to the
appointment of any new officer.

Que. What happens if there is a vacancy with either my chairman or
treasurer?\
Ans. If you are missing an officer you may not have any financial
activity until a new officer is appointed. It is very important for a
campaign finance entity to fill any vacancy either by resignation or
death immediately.

Que. Can a candidate join a slate committee without having a candidate
committee?\
Ans. No, candidates must have their own campaign finance entity
established prior to joinng a slate committee.

Que. Are candidates still members of a slate after the election?\
Ans. Only if the candidate wins election, otherwise the candidate will
be removed from the slate. A slate must have two active candidates as
member to remain active.

Que. My central committee appointed me Treasurer, do I notify SBE?\
Ans. Yes, if your party central committee voted a new officer, the new
officer has to send in the appointment form to SBE and the previous
officer needs to send SBE a resignation letter.

Bank Accounts\
Que. What is required to open a campaign depository account at a bank?\
Ans. Most banks will require a Form SS-4 (Application for Employer
Identification Number), personal identification and a Resolution of
Authority.

Que. What documents satisfy the Resolution of Authority requested by
banks to open a depository account?\
Ans. A bank should accept the processed Statement of Organization for a
newly formed committee by the SBE to open a depository account.

Using Campaign Finance Funds\
Que. I have a State campaign finance entity but want to run for U.S.
Congress, can I transfer over the monies to my federal campaign entity?\
Ans. Generally, Maryland campaign funds are for State candidates.
However, it is permissible for a State campaign finance entity to make
expenditures to support a federal candidate or federal PAC. In order to
make a contribution, the entity's account must have federally
permissible funds equaling the contribution. Please contact the Federal
Election Commission to find out what are the contribution limits and
reporting requirement, if any.

Que. Can a campaign finance entity pay late fees?\
Ans. Yes, Maryland law states that a campaign finance entity shall pay
for late fees. If the committee is unable to pay the late fees, then it
becomes the personal liability of the Chair and Treasurer.

Que. Can I invest campaign funds in a certificate of deposit (CD) that
matures before the election?\
Ans. Yes. However, there are limits on how campaign funds may be
invested. Please review COMAR 33.13.06.03B before making any
investments.

Campaign Finance\
Que. Are there limits to cash contributions?\
Ans. Yes, a person may make cash contributions up to \$100 for an
election cycle. Once the person contributes \$100, the person can no
longer make any cash contributions for the remainder of the election
cycle.

Que. Are campaign treasurers required to keep a copy of receipts issued
for campaign contributions?\
Ans. It is not required for treasurers to keep copies of contributions
receipts. The only receipt that the law requires treasurers to keep as
part of the treasurer's books and records is the receipt from a campaign
expenditure.

Que. If I do a "Pass the Hat" fundraiser, is it required I know all who
contributed and the amount?\
Ans. Yes. The committee needs to know the name and address of every
contributor and amount they gave, otherwise the income is anonymous and
illegal.

Que. Is the use of a business telephone(s) by a campaign for election
related activities such as polling an in-kind contribution?\
Ans. Yes. The campaign must report the use and cost of the telephone(s)
as an in-kind contribution subject to the \$6,000 contribution limit.

Que. For every credit card contribution, a fee is assessed for the
transaction. How should this be reported?\
Ans. A person makes a credit card contribution of \$100 to a candidate
committee. The candidate committee hired a service to process the
transaction. The hired service charges 1% on the amount per transaction.
The candidate committee reports receiving a contribution of \$100 from
the person and also reports making an expenditure of \$1 to the hired
service. It is not necessary to make the expenditures for every
transaction but the cumulative total weekly or monthly depending on how
the committee is billed.

General Questions\
Que. Can a federal PAC give to a Maryland campaign finance entity?\
Ans. Yes, a federal PAC is treated like a person under Maryland law. The
contribution limits are \$6,000 per campaign finance entity and \$24,000
aggregate for the election cycle.

Que. Can I use old campaign material for this campaign?\
Ans. Yes, but only if the committee name and authority line remains
identical. For example, a candidate committee account cannot use the
same campaign paraphernalia from a previous campaign if the treasurer is
different.

Closing Your Committee\
Que. Can I close my authorized candidate campaign committee while I'm
still on the ballot?\
Ans. No, the candidate committee must remain active and open during the
election. If you lose in the primary election, you may close your
committee with the first report due AFTER the Primary election. You may
not close your committee prior to the Primary election if you are an
active candidate on the Primary election ballot.

Independent Expenditures and Participating Organizations\
Que. If an out-of_State political committee makes independent
expenditures must it file independent expenditures reports?\
Ans. Yes. Even if the out-of-State political committee has registered
with SBE under Election Law Article 13-301, the out-of-State political
committee is not exempt from the disclosure obligations for making
independent expenditures.

Que. Does a federal political committee need to register as a
participating organization and file disbursement reports if it makes a
contribution to a Maryland candidate?\
Ans. Yes, If a federal political committee makes a contribution of
\$6,000 or more to a Maryland candidate, the federal committee needs to
register as a participating organization. If the federal committee makes
aggregate disbursements of \$10,000 or more in an election cycle, it is
required to file a report of all disbursements made to influence
elections in Maryland.

Making Expenditures\
Que. Can a campaign finance entity use a debit card to make cash
withdrawals to make expenditures?\
Ans. No, while it is permissible to use a debit card for purchases, it
is not permitted to make cash withdrawals from it.

Slates\
Que. Do central committee candidate have to establish a candidate
committee in order to join a slate?\
Ans. Yes. A central committee candidate would have to establish a
separate candidate campaign committee in order to join a slate. Pursuant
to EL 13-209, only after a establishing an authorized candidate campaign
committee may a candidate join a slate. Additionally, a contribution to
slate is considered a contribution to all of its members. If a candidate
for the central committee, then the candidate is required to establish
an authorized candidate committee. Therefore, a central committee
candidate that applies for the exemption of establishing a political
committee may not join or be member of the slate.
>>>>>>> 5aeece36d897bd6159c145766988cc7431298d32
